g---
title: "sales data EDA"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(reshape2)
library(ggplot2)
library(maps)
```

# Read, process data

```{r}
# load raw data sheet
sales <- read_excel("2017 Sales Summary.xlsx",
                    sheet = "Raw Data From Adroit")

# web sales data sheet
web <- read_excel("2017 Sales Summary.xlsx",
                  sheet = "STM Web Sales Summary",
                  col_names = F)
```

View before doing anything to raw sales data
```{r}
glimpse(sales)
```

```{r}
# filter data to remove completely NA rows
sales <- sales[rowSums(is.na(sales)) != ncol(sales),]
```

### Web data needs reformatting to be a dataframe

```{r}
web[is.na(web)] <- 0
names(web) <- web[2,] %>% unlist
names(web)[names(web)=="0"] <- unlist(web[1,])[names(web)=="0"]

web <- web[-(1:2),]
```

Date column needs to be cleaned up

```{r}
web$Date <- as.numeric(web$Date)
web <- web[!is.na(web$Date),]

web$Date <- as.Date(web$Date,
                    origin = "1899-12-30")
```

### Website-specific dataframe creation
```{r}
website <- web[names(web)[c(1, 63:70)]]

change <- names(website)[-1]
for(x in seq_along(change)){
  website[[change[x]]] <- as.numeric(website[[change[x]]])}
names(website)[names(website)=="TOTAL"] <- "Total Web Sales"
website <- website %>% filter(Date < "2017-07-03")
glimpse(website)
```

## Add columns and features to raw data

### Dates

Add columns for other intervals of each date

```{r}
sales$month_created <- as.Date(cut(sales$Created,
  breaks = "month"))

sales$week_created <- as.Date(cut(sales$Created,
  breaks = "week"))

sales$month_active <- as.Date(cut(sales$Active,
  breaks = "month"))


```

### 3x4 vs non column
```{r}
table(sales$Product)
```

```{r}
sales$three_four <- ifelse(grepl("3x4", sales$Product),
                          "3x4", "Regular")
sales$product_level <- ifelse(grepl("Deluxe", sales$Product),
                             "Deluxe",
                             ifelse(grepl("Choice", sales$Product),
                                    "Choice", 
                                    ifelse(grepl("Economy", sales$Product),
                                                 "Economy", "Standard")))
                                                     
```


### Count column for aggregation

for ggplot2

```{r}
sales$Count <- rep(1, nrow(sales))
```


### Mapping state names for plotting purposes

ggplot2 requires the long form name of a state but this dataset only has state acronyms. I'm merging the data to **state.csv** in order to make graphs later

```{r, message=FALSE, warning=FALSE}
state_mapping <- read_csv("states.csv")
names(state_mapping) <- c("region","State") # for merge
state_mapping$region <- tolower(state_mapping$region) #for merge
head(state_mapping)
```

```{r}
sales <- left_join(sales, state_mapping)

substr(sales$region, 1,1) <- toupper(substr(sales$region,1,1)) # nice
```


The state data is latitute and longitude information for states - meta data needed for plots

```{r, message=FALSE, warning=FALSE}
state_data <- ggplot2::map_data("state")
glimpse(state_data)
```

##### **Merge** 

```{r}
state_mapping <- left_join(state_data, state_mapping,  by = "region")
glimpse(state_mapping)
```

```{r}
state_sales <- merge(sales, state_mapping, sort=F)
```
```{r}
glimpse(sales)
```


# Exploration
### Missing values

**Percentage of values missing from each column**

```{r}
map_dbl(sales, function(x) sum(is.na(x))/nrow(sales)) %>% sort
```

### Unique values
**Number of unique values in each column**

```{r}
map_dbl(sales, function(x) length(unique(x))) %>% sort
```

## **Graphs**

### Distribution of sales Amount ($ shown for each sale/account)

```{r, message=FALSE, warning=FALSE}
ggplot(sales, aes(Amount)) + geom_histogram(binwidth=10)
```


### Number of sales by product
```{r}
ggplot(sales, aes(Product)) + geom_bar() + coord_flip()
```

### Sales volume by product (Amount)
```{r}

```


### Number of sales by benefit

```{r}
ggplot(sales, aes(Benefit)) + geom_bar() + coord_flip()
```

### **STM Web Sales data  - Jan-Jun 2017**


```{r, fig.height=15, warning=FALSE}
melted_website <- melt(website, id.vars = "Date")

web_series <- melted_website %>%
  filter(Date != "2017-01-31") %>%
  ggplot(aes(x=Date, y=value)) + geom_line() + geom_smooth() + facet_grid(variable~., scales = "free_y")
web_series
```

### Web Sales Y variable scatter plots
```{r}
website %>% 
  filter(Date != "2017-01-31") %>%
  ggplot(aes(Sessions, `Total Web Sales`)) + geom_point() + geom_smooth()
```

```{r}
website %>% 
  filter(Date != "2017-01-31") %>%
  ggplot(aes(Users, `Total Web Sales`)) + geom_point() + geom_smooth()
```

```{r}
website %>% 
  filter(Date != "2017-01-31") %>%
  ggplot(aes(Pageviews, `Total Web Sales`)) + geom_point() + geom_smooth()#
```

```{r}
website %>% 
  filter(Date != "2017-01-31") %>%
  ggplot(aes(`Pages/ Session`, `Total Web Sales`)) + geom_point() + geom_smooth()
```


## Sales by state

```{r}
sales$State %>% table %>% sort(decreasing=T) %>% data.frame
```


## Inactive accounts

This could be an area where we provide more value than just reporting. Maybe some of these inactivities/cancellations can be controlled/reduced?

These columns **Date Inactive** and **Inactive Reason** are usually both filled out together
```{r}
table(is.na(sales$`Date Inactive`) == is.na(sales$`Inactive Reason`))
```

### Top reasons for account inactivity

```{r}
table(sales$`Inactive Reason`) %>% sort(decreasing=T) %>% data.frame %>% head(10)
```

# **Candidate graphs for reporting**

This section contains graphs they can potentially be interested in based on the other sheets they created in the Excel file. 

* Sales by state
* Sales by month
* Sales by product
* Monthly, daily sales


## Sales ($) by month

**The data is from 12/2016 - 07/2017**

```{r}
sales %>%
  group_by(month_created, Product) %>%
  summarise(total_amount = sum(Amount, na.rm = T)) %>%
  ggplot(aes(x=month_created, y=total_amount, fill=Product)) + geom_bar(stat="identity")
```

## Sales ($) by week (more granular)

```{r}
week_sales <- sales %>%
  group_by(week_created, Product) %>%
  summarise(total_amount = sum(Amount, na.rm = T)) %>%
  ggplot(aes(x=week_created, y=total_amount, fill=Product)) + geom_bar(stat="identity")
week_sales
```

## Number of sales/accounts by week

```{r}
sales %>%
  group_by(week_created, Product) %>%
  summarise(total_count = sum(Count, na.rm = T)) %>%
  ggplot(aes(x=week_created, y=total_count, fill=Product)) + geom_bar(stat="identity")
```

## Sales ($) by state for June (not working yet)

```{r}
st <- state_sales %>%
  group_by(region, lat, long, group) %>%
  summarise(total_amount = sum(Amount, na.rm = T))

ggplot(st, aes(long, lat)) + geom_polygon(aes(group=group, fill=total_amount)) +
  geom_path(colour="grey50")
```

## States as a proportion of TOTAL sales/accounts over time 
```{r, message=FALSE, warning=FALSE}
ggplot(sales) + geom_bar(mapping=aes(as.character(month_created), fill = State), position = "fill") + theme(axis.text.x = element_text(angle = 45))
```

# Notes, impressions

## Source information

There appear to be **3 main sources** of data that are in the combined sales summary

```{r}
# get sheet names from excel file
readxl::excel_sheets("2017 Sales Summary.xlsx")
```

These 3 sources are **Adroit Sales Summary**, **Asst Agencies**, and **STM Web Sales Summary**. They get added together to calculate totals in the **Combined Sales Summary** sheet. I initially thought that the **Raw Data From Adroit** sheet populated the 3 main sources but some simple checking has shown me that the numbers are off - Specifically i found that there are less entries for plans in Texas in the raw data (244) compared to the combined sales summary (300).

### Web data outlier on 1/31/2017


Removed Web data outlier on 1/31/2017 to better see overall trend of web sales indicators. 

Here are the graphs with the outlier put back 

```{r, fig.height=10}
melted_website %>% 
  filter(variable %in% c("Date","Total Web Sales", "Sessions","Users","Pageviews")) %>%
  ggplot(aes(x=Date, y=value)) + geom_line() + geom_smooth() + facet_grid(variable~., scales = "free_y")
```

